---
title: "CAPM & Rolling Window Regression"
author: "Marleen Bünker, Denise Trinz, Eya Ben Romdhane & Franziska Meyndt"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: united
  html_notebook:
    toc: true
    number_sections: true
---
# Introduction
**Topic 1: CAPM and Rolling Window Regression**

You are the CFO of Mercedes AG and are interested in the **expected market return** of your company. You want to estimate it empirically using the CAPM model. Rather than relying on a single estimation approach, you conduct various sensitivity analyses. This includes in particular the application of a **Rolling Window Regression**, which illustrates the development of the expected market return over time.

To run the code, a local path for the working directory must be set. The folder should contain all files needed to load external data. In the second step, the following packages must be installed and loaded.
```{r setup, echo=TRUE, results='hide'}
# This chunk sets global settings for the entire document

# Set working directory – please adjust to your own path
knitr::opts_chunk$set(
  root.dir = "/Users/franziskameyndt/Desktop/Rolling Window Regression", # e.g. project folder
  echo = TRUE,      # Show code
  warning = FALSE,  # Suppress warnings
  message = FALSE   # Suppress package messages
)

# Install packages (only needed once)
# install.packages(c("tidyquant", "tidyverse", "broom", "slider", "ggplot2", "plotly", "lubridate", "htmltools", "shiny"))

# Load required packages
library(tidyverse)  # Data manipulation, visualization etc.
library(broom)      # Tidy model outputs
library(slider)     # Rolling window calculations
library(ggplot2)    # Create plots
library(plotly)     # Interactive visualizations
library(lubridate)  # Date handling
library(htmltools)  # HTML content e.g. in Shiny
library(shiny)      # Interactive web apps
library(zoo)  # Time series functions like as.yearmon (e.g. month extraction)
library(tidyquant)  # For tq_get() and other finance functions
```

# Load Data {.tabset}
After setting up the working directory and installing the required packages, the next data preparation step is to download the necessary data for the CAPM & Rolling Window Regression. The following parameters are defined:
- Ri: Stock return = MBG.DE Adjusted Close prices.
- Rm: Market return = DAX (German Stock Index) as the reference market.
- Rf: Risk-free rate = "Daily yield of the current five-year Federal notes", Bundesbank (see Appendix).

## Monthly Returns
Load data from Yahoo Finance, set parameters & time period:
```{r}
start_date <- "2020-01-01"
end_date <- "2024-12-31"
symbols <- c("MBG.DE", "^GDAXI")  # Mercedes & DAX
```
First, we define the time period for which we want to analyse the data – here from January 2020 to December 2024. We also specify the ticker symbols: MBG.DE stands for Mercedes, ^GDAXI for the German stock index DAX.

Load the monthly risk-free rate (see Appendix for source and further information):
```{r}
risk_free_monthly <- readRDS("risk_free_monthly_5 years.rds") %>%
  mutate(
    date_rf = ceiling_date(date, unit = "month") - days(1),  # Month-end
    yearmon = as.yearmon(date_rf)
  ) %>%
  select(yearmon, rf_xts.Open)  # Only necessary columns

head(risk_free_monthly)
```
The risk-free rate is essential for the CAPM model. We load a prepared file with risk-free interest rates and adjust the date to the last day of each month to make them more comparable to the returns.

Fetch data:
```{r}
stock_data <- tq_get(symbols, from = start_date, to = end_date)
```
Using the tq_get() function from the tidyquant package, we retrieve the historical price data for the defined symbols over the specified time period.

Calculate monthly returns with month-end dates:
```{r}
monthly_returns <- stock_data %>%
  group_by(symbol) %>%  # Group by stock (Mercedes, DAX)
  tq_transmute(
    select = adjusted,              # Use the Adjusted Close price
    mutate_fun = periodReturn,      # Function to calculate returns
    period = "monthly",             # Calculate returns on a monthly basis
    col_rename = "return",          # Name the column "return"
    type = "log"                    # Log returns for regressions
  )
```
For each stock, we now calculate returns on a monthly basis. We use logarithmic returns, as they offer certain mathematical advantages in financial analysis. The date is set to the last day of the month.

Convert to wide format & rename columns:
```{r}
returns_wide <- monthly_returns %>%
  pivot_wider(names_from = symbol, values_from = return) %>%  # Convert to wide format: one column per stock
  drop_na() %>%  # Remove rows with missing values
  rename(
    mercedes = `MBG.DE`,     # Rename for better readability
    market = `^GDAXI`        # DAX as market benchmark
  ) %>%
  mutate(
    date = ceiling_date(date, unit = "month") - days(1),  # Calculate month-end
    yearmon = as.yearmon(date)                            # Year-month format for later joins
  )
```
To more easily compare the returns of Mercedes and the market index, we reshape the data into wide format. We rename the columns for clarity, remove missing values, and set the date to the last day of the month again.

Merge with risk-free rate & calculate excess returns:
```{r}
returns_final_monthly <- returns_wide %>%
  left_join(risk_free_monthly, by = "yearmon") %>%  # Join monthly returns with the risk-free rate
  mutate(
    R_f = rf_xts.Open,                               # Rename column for better readability
    excess_mercedes = mercedes - R_f,                # Calculate Mercedes excess return
    excess_market = market - R_f                     # Calculate market (DAX) excess return
  ) %>%
  select(date, yearmon, mercedes, market, R_f, excess_mercedes, excess_market)  # Select relevant columns
```
We now merge the monthly returns with the risk-free rates using the common yearmon key. We then calculate the so-called excess returns, i.e. the excess returns of Mercedes and the market compared to the risk-free investment.

Dataframe check:
```{r}
head(returns_final_monthly)
```
Finally, we verify that the final dataset was created correctly by inspecting the first few rows.

## Daily Returns
In contrast to the monthly dataset, here we load daily price data and daily risk-free rates, calculate daily log returns, and determine the excess returns on a daily basis. The workflow remains structurally the same – the only difference is that aggregation and merging are done by date (date) rather than by month (yearmon).

Load data from Yahoo Finance, set parameters & time period:
```{r}
start_date <- "2020-01-01"
end_date <- "2024-12-31"
symbols <- c("MBG.DE", "^GDAXI")  # Mercedes & DAX
```
Load the daily risk-free rate:
```{r}
risk_free_daily <- readRDS("risk_free_daily_5 years.rds")
head(risk_free_daily)
```
Fetch data:
```{r}
stock_data <- tq_get(symbols, from = start_date, to = end_date)
```
Calculate daily returns:
```{r}
daily_returns <- stock_data %>%
  group_by(symbol) %>%
  tq_transmute(
    select = adjusted,
    mutate_fun = periodReturn,
    period = "daily",         # ← CHANGE: now daily
    type = "log",             
    col_rename = "return"
  )
```
Convert to wide format:
```{r}
returns_wide <- daily_returns %>%
  pivot_wider(names_from = symbol, values_from = return) %>%
  drop_na()

returns_wide <- returns_wide %>%
  rename(mercedes = `MBG.DE`,
         market = `^GDAXI`)
```
Merge with risk-free rate:
```{r}
returns_final_daily <- returns_wide %>%
  left_join(risk_free_daily, by = "date") %>%  # merge by date instead of yearmon
  mutate(
    R_f = risk_free,  # from the risk_free_daily table
    excess_mercedes = mercedes - R_f,
    excess_market = market - R_f
  )
head(returns_final_daily)
```
# Interactive Visualization with Shiny App

**IMPORTANT**
We have developed an interactive tool that allows users to dynamically analyse Beta and other risk factors. Users can flexibly select different time periods and window sizes to examine the evolution of the metrics over time in detail. The tool even offers extended time windows beyond the examples shown in the notebook. We warmly invite you to try the tool and explore the data for yourself.

**CLICK THE LINK:**
[Open Shiny App for Data Analysis](https://marleenb.shinyapps.io/shiny_app/)

# Exploratory Data Analysis

## Plot Returns as Time Series {.tabset}
Before starting the regression, we want to get an overview of the monthly & daily returns of Mercedes and the DAX. The visualizations allow us to identify first patterns, trends, or anomalies.

### Monthly Returns
```{r, echo=FALSE, results='asis'}
# Base plot with ggplot2
p_monthly <- ggplot(returns_final_monthly, aes(x = date)) +
  geom_line(aes(y = mercedes, color = "Mercedes")) +  # Line for Mercedes returns
  geom_line(aes(y = market, color = "DAX")) +         # Line for DAX returns
  labs(
    title = "Monthly Returns: Mercedes vs. DAX",  # Plot title
    y = "Return", x = "Date"                      # Axis labels
  ) +
  scale_color_manual(values = c("Mercedes" = "#A62321", "DAX" = "darkgrey")) +  # Assign colors
  scale_x_date(
    date_breaks = "3 months",  # x-axis in 3-month steps
    labels = function(x) paste0("Q", lubridate::quarter(x), " ", lubridate::year(x))  # Quarterly axis labels
  ) +
  theme_minimal() +  # Clean, minimalist design
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability

# Interactive display with Plotly in an HTML div (full width)
htmltools::browsable(
  htmltools::tags$div(
    style = "width:100%;",  # Set width to 100%
    ggplotly(p_monthly) %>%
      layout(autosize = TRUE) %>%     # Automatic resizing
      config(responsive = TRUE)       # Enable responsive behaviour
  )
)
```
Monthly returns:
- Due to the lower number of observations, estimates are more sensitive to outliers.
- Less data means higher variance in regression coefficients → Beta often only becomes significant at larger time windows.
- Fluctuations appear stronger, as monthly returns already bundle daily volatility.

### Daily Returns
```{r, echo=FALSE, results='asis'}
p_daily <- ggplot(returns_final_daily, aes(x = date)) +
  geom_line(aes(y = mercedes, color = "Mercedes")) +
  geom_line(aes(y = market, color = "DAX")) +
  labs(
    title = "Daily Returns: Mercedes vs. DAX",
    y = "Return", x = "Date"
  ) +
  scale_color_manual(values = c("Mercedes" = "#A62321", "DAX" = "darkgrey")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Wrapper div for full width
htmltools::browsable(
  htmltools::tags$div(
    style = "width:100%;",
    ggplotly(p_daily) %>%
      layout(autosize = TRUE) %>%
      config(responsive = TRUE)
  )
)
```
Daily returns:
- Individual daily returns show high short-term fluctuations, but are statistically more stable at the aggregate level.
- The large number of observations smooths out random outliers, so Beta estimates are often significant.
- Fine market signals and short-term reactions are captured more precisely.

## Scatterplot for Correlation {.tabset}
To visually and statistically examine the relationship between the excess return of Mercedes (excess_mercedes) and the market (excess_market), we create a scatterplot with a regression line and calculate the correlation coefficient.

### Monthly Returns
```{r}
# Base plot with scatter points
fig <- plot_ly(data = returns_final_monthly, x = ~excess_market, y = ~excess_mercedes,
               type = 'scatter', mode = 'markers',                    # Scatterplot with points
               marker = list(color = 'blue'), name = 'Data Points')   # Color and legend name

# Calculate regression line
regression <- lm(excess_mercedes ~ excess_market, data = returns_final_monthly)  # Fit linear model
x_vals <- seq(min(returns_final_monthly$excess_market, na.rm = TRUE),            # x values for line (min to max)
              max(returns_final_monthly$excess_market, na.rm = TRUE), length.out = 100)
y_vals <- predict(regression, newdata = data.frame(excess_market = x_vals))       # Predict y values for line

# Add regression line
fig <- fig %>% 
  add_trace(x = x_vals, y = y_vals, type = 'scatter', mode = 'lines',            # Add line
            line = list(color = 'red'), name = 'Regression Line') %>%             # Line in red with legend name
  layout(title = list(text = "Regression: excess_mercedes ~ excess_market", x = 0.5), # Centered title
         xaxis = list(title = "Excess Market Return"),                          # x-axis label
         yaxis = list(title = "Excess Mercedes Return"))                        # y-axis label

# Display plot at full width
browsable(
  tags$div(
    style = "width:100%;",                                                       # Container at 100% width
    fig %>% layout(autosize = TRUE) %>% config(responsive = TRUE)                # Make plot responsive
  )
)
```
Calculate correlation coefficient:
```{r}
cor(returns_final_monthly$mercedes, returns_final_monthly$market)
```
A value close to +1 indicates a strong positive correlation, while a value close to 0 indicates little or no linear relationship.

### Daily Returns
```{r}
# Base plot with scatter points for daily excess returns
fig <- plot_ly(data = returns_final_daily, x = ~excess_market, y = ~excess_mercedes,
               type = 'scatter', mode = 'markers',
               marker = list(color = 'blue'), name = 'Data Points')

# Calculate regression line
regression <- lm(excess_mercedes ~ excess_market, data = returns_final_daily)
x_vals <- seq(min(returns_final_daily$excess_market, na.rm = TRUE),
              max(returns_final_daily$excess_market, na.rm = TRUE), length.out = 100)
y_vals <- predict(regression, newdata = data.frame(excess_market = x_vals))

# Add regression line
fig <- fig %>% 
  add_trace(x = x_vals, y = y_vals, type = 'scatter', mode = 'lines',
            line = list(color = 'red'), name = 'Regression Line') %>%
  layout(title = list(text = "Regression (daily): excess_mercedes ~ excess_market", x = 0.5),
         xaxis = list(title = "Excess Market Return (daily)"),
         yaxis = list(title = "Excess Mercedes Return (daily)"))

# Display plot at full width
browsable(
  tags$div(
    style = "width:100%;",
    fig %>% layout(autosize = TRUE) %>% config(responsive = TRUE)
  )
)
```
Calculate correlation coefficient:
```{r}
cor(returns_final_daily$mercedes, returns_final_daily$market)
```
A value close to +1 indicates a strong positive correlation, while a value close to 0 indicates little or no linear relationship.

# CAPM
In the next step, we run a classical CAPM regression for the excess returns of Mercedes relative to the market. We first inspect the linear model summary, then present the results in a table, and finally visualize the relationship between market and Mercedes returns including the regression line in an interactive plot.

## Fit Linear Model {.tabset}

### Monthly Returns
```{r}
capm_model <- lm(excess_mercedes ~ excess_market, data = returns_final_monthly)
summary(capm_model)
```
Results as a table:
```{r}
tidy(capm_model)
```

### Daily Returns
```{r}
capm_model <- lm(excess_mercedes ~ excess_market, data = returns_final_daily)
summary(capm_model)
```
Results as a table:
```{r}
tidy(capm_model)
```
## Visualize CAPM Line {.tabset}
### Monthly Returns
```{r}
p_capm <- ggplot(returns_final_monthly, aes(x = excess_market, y = excess_mercedes)) +
  geom_point(alpha = 0.5, color = "#A62321", size = 2) +
  geom_smooth(method = "lm", color = "steelblue", se = TRUE, size = 1) +
  labs(
    title = "CAPM Estimation: Mercedes",
    x = "Excess Market Return",
    y = "Excess Mercedes Return"
  ) +
  theme_minimal(base_size = 12)

# Wrapper for full width
browsable(
  tags$div(
    style = "width:100%;",
    ggplotly(p_capm, tooltip = c("x", "y")) %>%
      layout(autosize = TRUE) %>%
      config(responsive = TRUE)
  )
)
```

### Daily Returns
```{r}
p_capm_daily <- ggplot(returns_final_daily, aes(x = excess_market, y = excess_mercedes)) +
  geom_point(alpha = 0.5, color = "#A62321", size = 2) +
  geom_smooth(method = "lm", color = "steelblue", se = TRUE, size = 1) +
  labs(
    title = "CAPM Estimation: Mercedes",
    x = "Excess Market Return",
    y = "Excess Mercedes Return"
  ) +
  theme_minimal(base_size = 12)

# Wrapper for full width
browsable(
  tags$div(
    style = "width:100%;",
    ggplotly(p_capm_daily, tooltip = c("x", "y")) %>%
      layout(autosize = TRUE) %>%
      config(responsive = TRUE)
  )
)
```
# Rolling Window Regression
The following section demonstrates how we use rolling regressions across various window sizes to analyse the temporal evolution of Beta coefficients for Mercedes relative to the market. For each time window, the regression coefficients including significance are extracted and stored in a list.

## Regression Calculation {.tabset}

### Monthly Returns
Rolling Regression:
```{r}
library(dplyr) # For data manipulation
library(slider) # For rolling windows
library(broom) # To extract model results
library(tibble) # Optional: for modern data structure (tibble instead of data.frame)

# Window sizes
window_sizes_monthly <- c(6, 12, 24)

# Store results as a list
rolling_results_monthly_list <- list()

# Loop
for (window_size in window_sizes_monthly) {
  rolling_results <- returns_final_monthly %>%
    drop_na(excess_mercedes, excess_market) %>% # Remove missing values
    slide_dfr(
      .f = ~ {
        # Rolling regression within the current window
        model <- lm(excess_mercedes ~ excess_market, data = .x)
        tidy_res <- tidy(model) %>% mutate(date = max(.x$date)) # Coefficients + date
        glance_res <- glance(model) %>% select(r.squared) # Optional: model fit (R²)
        bind_cols(tidy_res, glance_res) # Combine
      },
      .before = window_size - 1, # Set window size
      .complete = TRUE   # Use only complete windows
    ) %>%
    mutate(significant = if_else(p.value < 0.05, TRUE, FALSE)) # Flag significance
  
  rolling_results_monthly_list[[as.character(window_size)]] <- rolling_results
}
```
### Daily Returns
We apply the same approach at the daily return level using shorter window sizes due to the large number of observations.
Rolling Regression:
```{r}
# Window sizes for daily analysis (3, 6, 12 months ≈ 60, 120, 250 trading days)
window_sizes_daily <- c(60, 120, 250)

# Store results as a list
rolling_results_daily_list <- list()

for (window_size in window_sizes_daily) {
  rolling_results <- returns_final_daily %>%
    drop_na(excess_mercedes, excess_market) %>%
    slide_dfr(
      .f = ~ {
        model <- lm(excess_mercedes ~ excess_market, data = .x)
        tidy(model) %>% mutate(date = max(.x$date))
      },
      .before = window_size - 1,
      .complete = TRUE
    ) %>%
    filter(term == "excess_market") %>%
    mutate(significant = if_else(p.value < 0.05, TRUE, FALSE))
  
  rolling_results_daily_list[[as.character(window_size)]] <- rolling_results
}
```
## Visualize Rolling Beta {.tabset}
### Monthly Returns
In the next step, we visualize the Rolling Beta of Mercedes using an example window size of 24 months to analyse the temporal evolution of the risk metric relative to the market.

As a reference value, we use the Beta for the European automotive industry estimated by Damodaran. The Damodaran Beta value of 1.2 used in our analysis represents an approximation that lies between the unlevered Beta of 0.83 and the levered industry Beta of 1.63. This is intended to strike a balanced compromise between theoretical modelling and real-world practice. (Source: https://pages.stern.nyu.edu/~adamodar/New_Home_Page/datacurrent.html#discrate).
```{r}
# Filter only the Beta estimates (coefficients of excess_market) from the 24-month results
rolling_beta_monthly_24_beta_only <- rolling_results_monthly_list[["24"]] %>% 
  filter(term == "excess_market")

# Define plot using Beta-only data
p_monthly <- ggplot(rolling_beta_monthly_24_beta_only, aes(x = date)) +
  geom_line(aes(y = estimate, color = "Mercedes Beta"), size = 1) +        # Draw rolling Beta as a line
  geom_hline(aes(yintercept = 1.2, linetype = "Damodaran Beta"),           # Horizontal reference line (Damodaran)
             color = "grey40", size = 0.8) +
  scale_color_manual(name = NULL, values = c("Mercedes Beta" = "#A62321")) + # Define line colors
  scale_linetype_manual(name = NULL, values = c("Damodaran Beta" = "dashed")) + # Line type for Damodaran Beta
  labs(
    title = "Rolling Beta (Window = 24 Months) – Monthly Returns",         # Plot title
    x = NULL,                                                              # No x-axis label
    y = NULL                                                               # No y-axis label
  ) +
  theme_minimal() +                                                        # Use minimalist theme
  theme(
    legend.position = "bottom",                                            # Place legend at the bottom
    panel.grid = element_blank(),                                          # No grid lines
    axis.title.x = element_blank(),                                        # Remove x-axis title
    axis.title.y = element_blank(),                                        # Remove y-axis title
    plot.title = element_text(hjust = 0.5, face = "bold")                  # Center and bold title
  )

# Generate interactive output at full width
htmltools::browsable(
  htmltools::tags$div(
    style = "width:100%;",                                                  # Container at full width
    ggplotly(p_monthly) %>%                                                 # Convert ggplot to interactive Plotly
      layout(autosize = TRUE) %>%                                           # Automatic resizing
      config(responsive = TRUE)                                             # Enable responsive behaviour
  )
)
```
In the Beta visualization, we can see how the risk of Mercedes relative to the overall market (DAX) has changed over the past 24 months. The Rolling Beta shows fluctuations, meaning that the sensitivity of the Mercedes stock to market movements is not constant but adjusts over time.

A Beta above 1 indicates that Mercedes reacts more strongly to market fluctuations than the market itself, while a Beta below 1 suggests lower volatility compared to the market. The dashed horizontal line at 1.2 (Damodaran Beta) serves as a reference value for comparison with the estimated Betas. Overall, this visualization helps us better understand the dynamic risk profile of the stock and observe whether and when market risk has changed.

### Daily Returns
Here we use an example window size of 120 days (6 months) for the visualization.
```{r}
# Extract the 120-day window
rolling_beta_daily_120 <- rolling_results_daily_list[["120"]]

# Define plot
p_daily <- ggplot(rolling_beta_daily_120, aes(x = date)) +
  geom_line(aes(y = estimate, color = "Mercedes Beta"), size = 1) +
  geom_hline(aes(yintercept = 1.2, linetype = "Damodaran Beta"),
             color = "grey40", size = 0.8) +
  scale_color_manual(name = NULL, values = c("Mercedes Beta" = "#A62321")) +
  scale_linetype_manual(name = NULL, values = c("Damodaran Beta" = "dashed")) +
  labs(
    title = "Rolling Beta (Window = 120 Days) – Daily Returns",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

# Interactive output
htmltools::browsable(
  htmltools::tags$div(
    style = "width:100%;",
    ggplotly(p_daily) %>%
      layout(autosize = TRUE) %>%
      config(responsive = TRUE)
  )
)
```
## Significance Plot {.tabset}
The significance analysis of Beta values is central to making reliable statements about the systematic risk of a company relative to the overall market. Beta values feed into the calculation of the cost of equity, company valuations, and investment and financing decisions.

However, a Beta value is only meaningful when it is **statistically significant**. A non-significant Beta value means that the observed relationship with the market could be random. The estimate is then uncertain and should not be used as a basis for decision-making.

### Monthly Returns
Monthly vs. Daily:
For a sound decision-making basis, it is more appropriate to focus on monthly returns, because:
- They are less affected by short-term noise.
- Beta estimates become more stable and meaningful over longer time periods.
- They provide a better assessment of long-term systematic risk.

While daily data can be useful for short-term analyses, it should not serve as the primary basis for strategic decisions.
```{r, echo=FALSE}
# Initialize empty list for Plotly traces
plot_data_monthly <- list()
i <- 1

# Loop over different window sizes (window_sizes_monthly)
for (window_size in window_sizes_monthly) {
  # Load rolling results for current window
  rolling_results <- rolling_results_monthly_list[[as.character(window_size)]]

  # Filter only Beta coefficients (term == "excess_market")
  beta_results <- rolling_results %>% filter(term == "excess_market")

  # Significant Beta values as scatter points (green circles)
  trace_sig <- list(
    x = beta_results$date[beta_results$significant],       # Dates for significant values only
    y = beta_results$estimate[beta_results$significant],   # Beta estimates (significant)
    type = 'scatter',                                       # Scatterplot
    mode = 'markers',                                       # Points only
    name = paste0("Significant (", window_size, "M)"),     # Legend name with window size
    visible = ifelse(i == 1, TRUE, FALSE),                  # Visible only for first window by default
    marker = list(color = "darkgreen", size = 6, symbol = "circle")  # Marker style
  )

  # Non-significant Beta values as scatter points (red crosses)
  trace_nonsig <- list(
    x = beta_results$date[!beta_results$significant],      # Dates of non-significant values
    y = beta_results$estimate[!beta_results$significant],  # Beta estimates (non-significant)
    type = 'scatter',
    mode = 'markers',
    name = paste0("Not significant (", window_size, "M)"),
    visible = ifelse(i == 1, TRUE, FALSE),                  # Visible only for first window
    marker = list(color = "red", size = 6, symbol = "x")    # Marker style (red cross)
  )

  # Line for Beta trend over time
  trace_line <- list(
    x = beta_results$date,                                  # Date
    y = beta_results$estimate,                              # Beta estimates
    type = 'scatter',
    mode = 'lines',                                         # Line mode
    name = paste0("Beta ", window_size, "M"),              # Legend name with window size
    visible = ifelse(i == 1, TRUE, FALSE),                  # Visible only for first window
    line = list(color = "#2C3E50", width = 2)               # Line color and width
  )

  # Add traces to the list: line, significant points, non-significant points
  plot_data_monthly[[length(plot_data_monthly) + 1]] <- trace_line
  plot_data_monthly[[length(plot_data_monthly) + 1]] <- trace_sig
  plot_data_monthly[[length(plot_data_monthly) + 1]] <- trace_nonsig

  # Increment index for next window
  i <- i + 1
}

# Number of traces per window size (line + significant + not significant)
num_traces_per_window <- 3

# Prepare buttons for selecting window size
buttons_monthly <- lapply(seq_along(window_sizes_monthly), function(i) {
  # Set all traces to invisible initially
  visibility <- rep(FALSE, length(plot_data_monthly))
  
  # Make only the three traces of the current window visible
  start <- (i - 1) * num_traces_per_window + 1
  visibility[start:(start + 2)] <- TRUE

  # Button definition
  list(
    method = "update",                         # Update visibility on click
    args = list(list(visible = visibility)), # Control trace visibility
    label = paste0(window_sizes_monthly[i], " Months")  # Button label with window size
  )
})

# Initialize empty Plotly object
fig_monthly <- plot_ly()

# Add all prepared traces to the plot
for (trace in plot_data_monthly) {
  fig_monthly <- fig_monthly %>% add_trace(
    x = trace$x,
    y = trace$y,
    type = trace$type,
    mode = trace$mode,
    name = trace$name,
    visible = trace$visible,
    line = trace$line,
    marker = trace$marker
  )
}

# Define plot layout with title, axis labels, and update buttons
fig_monthly <- fig_monthly %>%
  layout(
    title = list(text = "Rolling Beta – Monthly Returns with Significance", x = 0.5),
    xaxis = list(title = "Date"),
    yaxis = list(title = "Beta"),
    updatemenus = list(
      list(
        type = "buttons",             # Buttons for switching
        direction = "down",           # Vertical button arrangement
        x = 1.05,                    # Position to the right of the plot
        xanchor = "left",
        y = 0.8,
        yanchor = "top",
        pad = list(r = 10, t = 10),  # Padding from plot border
        buttons = buttons_monthly,    # List of buttons
        showactive = TRUE             # Highlight active button
      )
    ),
    legend = list(orientation = "h", x = 0.1, y = -0.25),  # Horizontal legend below plot
    margin = list(r = 120)  # Extra space on the right for buttons
  )

# Output plot as HTML widget at full width
htmltools::tags$div(
  style = "width: 100%;",
  plotly::ggplotly(fig_monthly)
)
```
When calculating Beta based on monthly returns, a statistically significant relationship only emerges from a time window of approximately 24 months. This can be explained as follows:
- **Small sample size**: Short time windows (e.g. 6–12 months) provide too few data points to enable stable estimates.
- **High standard errors:** Uncertain estimates lead to higher p-values and thus to a lack of significance.

Only with a larger number of observations over a longer time period can Beta be estimated more precisely and its significance statistically confirmed.

### Daily Returns
Monthly vs. Daily:
For a sound decision-making basis, it is more appropriate to focus on monthly returns, because:
- They are less affected by short-term noise.
- Beta estimates become more stable and meaningful over longer time periods.
- They provide a better assessment of long-term systematic risk.

While daily data can be useful for short-term analyses, it should not serve as the primary basis for strategic decisions.
```{r, echo=FALSE}
# Plotly data list for daily
plot_data_daily <- list()
i <- 1

for (window_size in window_sizes_daily) {
  rolling_results <- rolling_results_daily_list[[as.character(window_size)]]

  # Only Beta coefficients for significance
  beta_results <- rolling_results %>% filter(term == "excess_market")

  # Significant values (Beta only)
  trace_sig <- list(
    x = beta_results$date[beta_results$significant],
    y = beta_results$estimate[beta_results$significant],
    type = 'scatter',
    mode = 'markers',
    name = paste0("Significant (", window_size, "D)"),
    visible = ifelse(i == 1, TRUE, FALSE),
    marker = list(color = "darkgreen", size = 6, symbol = "circle")
  )

  # Non-significant values (Beta only)
  trace_nonsig <- list(
    x = beta_results$date[!beta_results$significant],
    y = beta_results$estimate[!beta_results$significant],
    type = 'scatter',
    mode = 'markers',
    name = paste0("Not significant (", window_size, "D)"),
    visible = ifelse(i == 1, TRUE, FALSE),
    marker = list(color = "red", size = 6, symbol = "x")
  )

  # Line for Beta trend
  trace_line <- list(
    x = beta_results$date,
    y = beta_results$estimate,
    type = 'scatter',
    mode = 'lines',
    name = paste0("Beta ", window_size, "D"),
    visible = ifelse(i == 1, TRUE, FALSE),
    line = list(color = "#2C3E50", width = 2)
  )

  plot_data_daily[[length(plot_data_daily) + 1]] <- trace_line
  plot_data_daily[[length(plot_data_daily) + 1]] <- trace_sig
  plot_data_daily[[length(plot_data_daily) + 1]] <- trace_nonsig

  i <- i + 1
}

# Buttons for daily
num_traces_per_window_daily <- 3
buttons_daily <- lapply(seq_along(window_sizes_daily), function(i) {
  visibility <- rep(FALSE, length(plot_data_daily))
  start <- (i - 1) * num_traces_per_window_daily + 1
  visibility[start:(start + 2)] <- TRUE

  list(
    method = "update",
    args = list(list(visible = visibility)),
    label = paste0(window_sizes_daily[i], " Days")
  )
})

# Final plot for daily
fig_daily <- plot_ly()

for (trace in plot_data_daily) {
  fig_daily <- fig_daily %>% add_trace(
    x = trace$x,
    y = trace$y,
    type = trace$type,
    mode = trace$mode,
    name = trace$name,
    visible = trace$visible,
    line = trace$line,
    marker = trace$marker
  )
}

fig_daily <- fig_daily %>%
  layout(
    title = list(text = "Rolling Beta – Daily Returns with Significance", x = 0.5),
    xaxis = list(title = "Date"),
    yaxis = list(title = "Beta"),
    updatemenus = list(
      list(
        type = "buttons",
        direction = "down",
        x = 1.05,
        xanchor = "left",
        y = 0.8,
        yanchor = "top",
        pad = list(r = 10, t = 10),
        buttons = buttons_daily,
        showactive = TRUE
      )
    ),
    legend = list(orientation = "h", x = 0.1, y = -0.25),
    margin = list(r = 120),
    autosize = TRUE
  ) %>%
  config(responsive = TRUE)

# Display plot
htmltools::tags$div(
  style = "width: 100%;",
  plotly::ggplotly(fig_daily)
)
```
Beta values for daily returns and shorter time windows appear significant because:
- The large number of daily data points increases statistical power.

However, caution is advised, as:
- Short-term fluctuations (noise) can lead to biased but nonetheless significant results.
- Significance does not necessarily reflect a reliable long-term risk assessment.

# Appendix
In addition to the Rolling Beta analysis, we examine the monthly Rolling Volatility, Rolling Sharpe Ratio, and Rolling Alpha to obtain a more comprehensive picture of Mercedes' risk, return, and performance dynamics over time. Caution is warranted, however, particularly with regard to Rolling Alpha, where the significance of the estimates should be carefully considered to avoid misinterpretation.

## Rolling Alpha
**Definition:** Alpha measures the risk-adjusted excess return of a portfolio relative to a benchmark and indicates whether an active manager has created additional value.
**What it shows:** The time-varying excess return of a portfolio compared to a benchmark (e.g. market portfolio).
**What can be observed:** Whether the portfolio has outperformed or underperformed the market over specific time periods – i.e. whether it consistently adds value.
```{r}
## Extract Rolling Alpha
rolling_alpha_24 <- rolling_results_monthly_list[["24"]] %>%
  filter(term == "(Intercept)") %>%
  mutate(date = as.Date(date)) %>%
  mutate(
    t_value = estimate / std.error,
    p_value = 2 * pt(-abs(t_value), df = 24 - 2),
    signif = ifelse(p_value < 0.05, "significant", "not significant")
  )

# Plot with line and points, color points by significance
p_alpha_24 <- ggplot(rolling_alpha_24, aes(x = date, y = estimate)) +
  geom_line(linewidth = 0.5, color = "gray40") +  # Line in neutral grey
  geom_point(aes(color = signif), size = 1) +   # Points colored by significance
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +  # Zero line
  scale_color_manual(values = c("significant" = "darkgreen", "not significant" = "red")) +
  labs(
    title = "Rolling Alpha of Mercedes-Benz AG (24 Months)",
    x = "Date",
    y = "Alpha",
    color = "Significance"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.grid = element_blank()
  )

# Interactive output with Plotly
htmltools::tags$div(
  style = "width: 100%;",
  plotly::ggplotly(p_alpha_24)
)
```
## Rolling R²
**Definition:** The coefficient of determination (R²) indicates how well a regression model explains the variance of the dependent variable, and in financial analysis often reflects the degree of market dependence of a portfolio.
**What it shows:** The proportion of the variance in portfolio returns explained by the market (benchmark), calculated over rolling time windows.
**What can be observed:** How strongly the portfolio depends on the market during different periods. A declining R² may indicate reduced market dependence or a change in strategy.
```{r}
rolling_r2_tbl <- rolling_results_monthly_list[["24"]] %>%
  distinct(date, r.squared)

p_r2 <- plot_ly(
  data = rolling_r2_tbl,
  x = ~date,
  y = ~r.squared,
  type = 'scatter',
  mode = 'lines',
  line = list(color = 'purple'),
  name = "Rolling R²"
) %>%
  layout(
    title = "Rolling R² (Mercedes vs. Market, 24 Months)",
    xaxis = list(title = "Date"),
    yaxis = list(title = "R²", range = c(0, 1))
  )

htmltools::tags$div(
  style = "width: 100%;",
  p_r2
)
```
## Visualize Rolling Volatility & Rolling Sharpe Ratio
*Rolling Volatility*

**Definition:** Volatility is a measure of portfolio risk and describes how strongly returns fluctuate around their mean.
**What it shows:** The intensity of fluctuation (standard deviation) of portfolio returns over rolling time windows.
**What can be observed:** How stable or volatile the portfolio was in different periods. Higher volatility means more risk, lower volatility means more stability.

*Rolling Sharpe Ratio*

**Definition:** The Sharpe Ratio measures the excess return of a portfolio per unit of risk taken (volatility).
**What it shows:** The risk-adjusted return of a portfolio over time, calculated over rolling time windows.
**What can be observed:** Whether the portfolio was adequately compensated for the risk taken over time. Fluctuations indicate changes in risk or performance.
```{r}
# Calculate Rolling Volatility & Sharpe Ratio (annualized)
returns_final_monthly <- returns_final_monthly %>%
  arrange(date) %>%
  mutate(
    rolling_volatility_24 = slide_dbl(
      excess_mercedes,
      ~ sd(.x, na.rm = TRUE) * sqrt(12),
      .before = 23,
      .complete = TRUE
    ),
    rolling_sharpe_24 = slide_dbl(
      excess_mercedes,
      ~ {
        mean_excess <- mean(.x, na.rm = TRUE)
        sd_excess <- sd(.x, na.rm = TRUE)
        if (sd_excess == 0) return(NA_real_)
        (mean_excess / sd_excess) * sqrt(12)
      },
      .before = 23,
      .complete = TRUE
    )
  )

# Individual plots
p_vol <- plot_ly(returns_final_monthly, x = ~date, y = ~rolling_volatility_24, type = 'scatter', mode = 'lines',
                 line = list(color = 'blue'), name = "Rolling Volatility (24 Months)") %>%
  layout(yaxis = list(title = "Volatility"))

p_sharpe <- plot_ly(returns_final_monthly, x = ~date, y = ~rolling_sharpe_24, type = 'scatter', mode = 'lines',
                    line = list(color = 'green'), name = "Rolling Sharpe Ratio (24 Months)") %>%
  layout(yaxis = list(title = "Sharpe Ratio"))

# Create subplot
combined_plot <- subplot(p_vol, p_sharpe, nrows = 2, shareX = TRUE, titleX = TRUE) %>%
  layout(margin = list(l = 50, r = 50, t = 50, b = 50))

# HTML container with CSS for 100% width
htmltools::tags$div(
  style = "width: 100%; height: auto; min-height: 400px; max-height: 70vh;",
  combined_plot
)
```

## Additional Data Preparation
### Data Check
We conduct a data check of the loaded financial data by comparing the data from the tidyquant function with Yahoo Finance data for MBG.DE and the DAX imported via .csv. We follow the approach from Chapter 1 of the book:
Ang, C. S. (2021). Analyzing Financial Data and Implementing Financial Models Using R (2nd ed.). Springer Nature Switzerland AG. https://doi.org/10.1007/978-3-030-64155-9.

Load and clean CSV data:
```{r}
setwd("~/Desktop/Rolling Window Regression")

# MBG
mbg <- read.csv("MBG_DE_2020_2024.csv", skip = 2)
colnames(mbg)[1] <- "Date"
colnames(mbg) <- c("Date", "Adj_Close", "Close", "High", "Low", "Open", "Volume")
mbg$Date <- as.Date(mbg$Date)
numeric_cols <- c("Adj_Close", "Close", "High", "Low", "Open", "Volume")
mbg[numeric_cols] <- lapply(mbg[numeric_cols], as.numeric)
mbg_xts <- xts(mbg[, numeric_cols], order.by = mbg$Date)

# DAX
dax_csv <- read.csv("GDAXI_2020_2024.csv", skip = 2)
colnames(dax_csv)[1] <- "Date"
colnames(dax_csv) <- c("Date", "Adj_Close", "Close", "High", "Low", "Open", "Volume")
dax_csv$Date <- as.Date(dax_csv$Date)
numeric_cols <- c("Adj_Close", "Close", "High", "Low", "Open", "Volume")  # define again
dax_csv[numeric_cols] <- lapply(dax_csv[numeric_cols], as.numeric)
dax_xts <- xts(dax_csv[, "Adj_Close"], order.by = dax_csv$Date)
```

Price visualization:
```{r}
plot(mbg_xts$Adj_Close, main = "MBG Adj_Close Prices", ylab = "Adj. Close Price", xlab = "Date", col = "black")
```

Calculate German trading days:
```{r}
library(timeDate)
start_date <- as.Date("2020-01-01")
end_date <- as.Date("2024-12-31")
all_days <- seq(start_date, end_date, by = "days")
trading_days <- all_days[!weekdays(all_days) %in% c("Saturday", "Sunday")]

holidays_de <- c(
  as.Date(Easter(2020)) + c(-2, 1),
  as.Date(Easter(2021)) + c(-2, 1),
  as.Date(Easter(2022)) + c(-2, 1),
  as.Date(Easter(2023)) + c(-2, 1),
  as.Date(Easter(2024)) + c(-2, 1),
  as.Date(paste0(2020:2024, "-01-01")),
  as.Date(paste0(2020:2024, "-05-01")),
  as.Date(paste0(2020:2024, "-10-03")),
  as.Date(paste0(2020:2024, "-12-24")),
  as.Date(paste0(2020:2024, "-12-25")),
  as.Date(paste0(2020:2024, "-12-26")),
  as.Date(paste0(2020:2024, "-12-31"))
)

trading_days <- setdiff(trading_days, holidays_de)
length(trading_days)
```

Summary Statistics:
```{r}
summary(mbg_xts$Adj_Close,dax_xts$Adj_Close)
```

Load and normalize Yahoo Finance data:
```{r}
mbg_yahoo <- tq_get("MBG.DE", from = "2020-01-01", to = "2024-12-31")
dax_yahoo <- tq_get("^GDAXI", from = "2020-01-01", to = "2024-12-31")

mbg_yahoo_xts <- xts(mbg_yahoo$adjusted, order.by = mbg_yahoo$date)
dax_yahoo_xts <- xts(dax_yahoo$adjusted, order.by = dax_yahoo$date)

mbg_norm_yahoo <- (mbg_yahoo_xts / as.numeric(mbg_yahoo_xts[1])) * 100
dax_norm_yahoo <- (dax_yahoo_xts / as.numeric(dax_yahoo_xts[1])) * 100
```

Compare MBG CSV vs. Yahoo Finance:
```{r}
mbg_xts_adj <- mbg_xts$Adj_Close
mbg_norm_csv <- (mbg_xts_adj / as.numeric(mbg_xts_adj[1])) * 100

mbg_compare <- merge(mbg_norm_csv, mbg_norm_yahoo, all = TRUE)
colnames(mbg_compare) <- c("CSV", "Yahoo")
mbg_compare_clean <- na.omit(mbg_compare)

plot.zoo(mbg_compare_clean, plot.type = "single", col = c("blue", "green"),
         main = "Normalized Prices: MBG (CSV) vs. Yahoo Finance", ylab = "Normalized Price", xlab = "Date")
legend("topleft", legend = c("MBG (CSV)", "MBG (Yahoo)"), col = c("blue", "green"), lty = 1)
```

```{r}
library(kableExtra)
mbg_table <- data.frame(
  Date = index(mbg_compare_clean),
  MBG_CSV = as.numeric(mbg_compare_clean$CSV),
  MBG_Yahoo = as.numeric(mbg_compare_clean$Yahoo)
)
mbg_table$Difference <- abs(mbg_table$MBG_CSV - mbg_table$MBG_Yahoo)
mbg_differences <- mbg_table %>% filter(Difference > 0.01)

mbg_differences %>%
  head(10) %>%
  kable(digits = 4, caption = "Rows with Differences Between MBG (CSV) and Yahoo") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped"))
```

Compare DAX CSV vs. Yahoo Finance:
```{r}
dax_norm_csv <- (dax_xts / as.numeric(dax_xts[1])) * 100
dax_compare <- merge(dax_norm_csv, dax_norm_yahoo, join = "inner")
colnames(dax_compare) <- c("CSV", "Yahoo")

dax_table <- data.frame(
  Date = index(dax_compare),
  DAX_CSV = as.numeric(dax_compare[, "CSV"]),
  DAX_Yahoo = as.numeric(dax_compare[, "Yahoo"])
)
dax_table$Difference <- abs(dax_table$DAX_CSV - dax_table$DAX_Yahoo)
dax_differences <- dax_table %>% filter(Difference > 0.01)

dax_differences %>%
  head(10) %>%
  kable(digits = 4, caption = "Rows with Differences Between DAX (CSV) and Yahoo") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped"))
```

### Risk Free Rate {.tabset}
This section shows the individual steps for extracting the risk-free rate from an external R script. The source used is the daily yields of the current five-year Federal notes ("Daily yield of the current five-year Federal notes").
Source: https://www.bundesbank.de/en/statistics/money-and-capital-markets/interest-rates-and-yields/daily-yields-of-current-federal-securities-772220.

#### Monthly
```{r}
# Load packages
library(xts)
library(dplyr)
library(lubridate)
library(zoo)

# Set working directory
setwd("~/Desktop/Rolling Window Regression")

# Read CSV data (Federal bond)
rf <- read.csv("BBSSY.D.REN.EUR.A620.000000WT0505.A.csv", skip = 8)
colnames(rf) <- c("Date", "Bundesanleihe", "na")
rf <- rf %>% 
  select(-na) %>% 
  filter(Bundesanleihe != ".")

# Convert date format & numeric values
rf$Date <- as.Date(rf$Date)
rf$Bundesanleihe <- as.numeric(as.character(rf$Bundesanleihe))

# Filter to time period
rf <- rf %>% filter(Date >= as.Date("2020-01-01") & Date <= as.Date("2024-12-31"))

# Create xts object
rf_xts <- xts(rf$Bundesanleihe, order.by = rf$Date)
colnames(rf_xts) <- "Bundesanleihe"

# Aggregate to monthly data
rf_monthly_xts <- to.monthly(rf_xts)

# Calculate monthly return from Open value (annual rate)
monthly_rf_xts <- (1 + rf_monthly_xts[, "rf_xts.Open"] / 100)^(1 / 12) - 1

# Convert to dataframe
risk_free_monthly <- data.frame(
  date = index(monthly_rf_xts),
  risk_free = coredata(monthly_rf_xts)
)

# Convert yearmon → Date (month-end)
risk_free_monthly$date <- as.Date(as.yearmon(risk_free_monthly$date), frac = 1)

# Set date to month-end for CAPM alignment
risk_free_monthly <- risk_free_monthly %>%
  mutate(date = ceiling_date(date, unit = "month") - days(1))

# Preview
head(risk_free_monthly)
nrow(risk_free_monthly)  # Should equal the number of months

# Save for later import
saveRDS(risk_free_monthly, file = "risk_free_monthly_5 years.rds")
```
#### Daily
```{r}
# Load packages
library(xts)
library(dplyr)
library(lubridate)
library(zoo)

# Set working directory
setwd("~/Desktop/Rolling Window Regression")

# Read CSV data (Federal bond)
rf <- read.csv("BBSSY.D.REN.EUR.A620.000000WT0505.A.csv", skip = 8)
colnames(rf) <- c("Date", "Bundesanleihe", "na")

# Clean and format
rf <- rf %>% 
  select(-na) %>% 
  filter(Bundesanleihe != ".") %>%
  mutate(
    Date = as.Date(Date),
    Bundesanleihe = as.numeric(as.character(Bundesanleihe))
  ) %>%
  filter(Date >= as.Date("2020-01-01") & Date <= as.Date("2024-12-31"))

# Calculate daily risk-free rate (annualized → daily)
rf <- rf %>%
  mutate(
    risk_free = (1 + Bundesanleihe / 100)^(1 / 252) - 1  # 252 trading days
  ) %>%
  select(date = Date, risk_free)

# Preview
head(rf)
nrow(rf)

# Save
saveRDS(rf, file = "risk_free_daily_5 years.rds")
```
